\propertitle

### 2019-11-25 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-21T16:46:09+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Abstract
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abstractrunin
\abslabeldelim{}
\renewcommand*{\abstractname}{}
\setlength{\absleftindent}{0pt}
\setlength{\absrightindent}{0pt}
\setlength{\abstitleskip}{-\absparindent}
\begin{abstract}\labelsep 0pt%
  \noindent \mynote{Abstract must be rewritten once paper is ready}This
  work shows how to build a maximum-entropy probabilistic model for the
  total activity of a network of neurons, given only some activity data or
  statistics -- for example, empirical moments -- of a \emph{subnetwork}
  thereof. This kind of model is useful because neuronal recordings are
  always limited to a very small sample of a network of neurons.
%
  The model is applied to two sets of neuronal data available in the
  literature. In some cases it makes interesting forecasts about the larger
  network -- for example, two low-regime modes in the frequency
  distribution for the total activity -- that are not visible in the sample
  data or in maximum-entropy models applied only to the sample.
%
  For the two datasets, the maximum-entropy probability model applied only
  to the subnetwork is compared with the marginal probability
  distribution obtained from the maximum-entropy model applied to the full
  network. On a linear probability scale no large differences are
  visible, but on a logarithmic scale the two distributions show very
  different behaviours, especially in the tails.
  \\\noindent\emph{\footnotesize Note: Dear Reader \amp\ Peer, this
    manuscript is being peer-reviewed by you. Thank you.}
% \par%\\[\jot]
% \noindent
% {\footnotesize PACS: ***}\qquad%
% {\footnotesize MSC: ***}%
%\qquad{\footnotesize Keywords: ***}
\end{abstract}
\selectlanguage{british}\frenchspacing

@@

\section{Introduction:\\a simple model for questions about large neuronal
  networks}
\label{sec:intro}

@@

Despite remarkable advances in recording technologies, the best
experimental measurements of instantaneous neuronal activity can still only
record a very small sample of neurons -- hundreds at most -- compared to
the numbers that constitute a functionally distinguished brain region. Many
probabilistic models focus on such samples only, somehow neglecting, in
their assumptions, that the recorded neurons are a sample from a larger
network. This kind of isolation assumptions sometimes escape attention,
being subtly hidden in the mathematics. Some probabilistic models try to
take also unrecorded neurons into account individually, and therefore
become very complex\mynote{Yasser \amp\ Vahid: refs for here?}. It would be
useful to have models that operate in between: addressing the larger brain
area whence the sample comes, but without adding too much complexity and
detail about it. Such intermediate models would be useful, for example, for
preliminary investigations, to help us to decide which hypotheses to
discard and which to consider for more complex and costly studies.

\mynote{[Yasser:] In the present work we aim to address the issue of building probability distributions over a large network using recording from subnetworks by considering the distribution of total population activity as an example. [}
In the present work we propose such an intermediate model. It answers this
question:\mynote{]} How much can the \emph{total} activity of a large neuronal
network have been, given the observation of the activity of a very small
sample thereof? This model addresses a larger brain area, avoiding the
assumption of isolation of the sample; and by focusing on the total
activity, rather than the activity of individual neurons, it remains simple
and numerically tractable.

@@

We illustrate possible uses of our proposed model in
\sect~\ref{sec:application}, by applying it to two concrete data sets:
\begin{enumerate*}[label=(\alph*)]
\item the activity of 65 neurons recorded for 20 min from a rat's Medial
  Entorhinal Cortex \citep{stensolaetal2012}, \item the activity of 159
  neurons recorded for 15 min from a macaque's Motor Cortex
  \citep{rostamietal2016_r2017}.
\end{enumerate*}
For each data set the model gives the most plausible frequency distribution
with which all levels of total activity of a larger network of 1\,000--10\,000
neurons appeared during the recording. These guessed frequency
distributions are distinctly different from the recorded ones for the
sampled neurons. For the first data set, for example, the frequency
distribution for the full network has two very distinct modes, both at low
activities, %($0.7\%$ and $2.6\%$ of simultaneously active neurons
% within $3\;\mathrm{ms}$,
see \fig~\ref{fig:plots_5mom}, whereas the frequency distribution for the
sample is monotonically decreasing with its maximum at zero activity. The
frequency distribution for the second data set doesn't have two modes but
still presents one prominent shoulder in its low-activity mode. Note that
these guessed features of the full network cannot be inferred by the
application of maximum-entropy \emph{to the sample alone}. Although we don't
formulate any hypotheses on why the larger network could have two distinct
low-activity regimes -- the purpose of the present work is only
methodological -- these results shows that the proposed method can lead to
the formulation or preliminary assessment of interesting hypotheses.

We want to stress the usefulness of making a quantified guess about the
activity of a larger brain area. Such a guess seems indeed to be the
primary idea behind recording a sample from that area. %  Second, it allows us
% to make comparisons across experimental sessions; such comparisons would me
% difficult or meaningless if made with the recorded samples, which generally
% comprise non-overlapping sets of neurons and differ in size.
There's also an important methodological reason. Maximum-entropy methods
have been used to assess the informational sufficiency of pairwise
correlations and correlations of higher order \citep[see for
example][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}.
But when the difference in size between sample and full network is too
large, correlation sufficiency for the sample implies the \emph{lack} of
correlation sufficiency for the larger network, and vice versa. Therefore,
maximum-entropy applications at the sample level can be deceptive for
questions of statistical sufficiency, and a full-network application is
more reliable. We discuss this point in detail in \sect~\ref{sec:marginalization}.

@@

How large is the full network addressed by the method proposed here? Its
size must have some limit and can't obviously include the full brain. The
size is determined by the validity of the formulae from sampling theory and
discussed in \sect~\ref{sec:N}.

We obviously don't know whether the frequency distribution obtained with
our approach is the \emph{actual} one which the activity levels of the full
network had during the recording; it's only the most plausible. In high
dimensions, however, the features of the most plausible distribution may
\emph{not} be typical of the majority of most plausible distributions; and
the set of all possible frequency distributions, if the full network for
example comprises 1\,000 neurons, is a 1\,000-dimensional space. In
\sect~\ref{sec:assumptions} we therefore try to assess which features of
the frequency distribution delivered by our method may be typical and
therefore expected of the actual one. We find that general features such as
the bimodality of the first data set are indeed typical. Maximum-entropy
models can be considered as approximations of Bayesian models based on
various assumptions of inferential sufficiency
\citep{jaynes1986d_r1996,portamana2017}. How do our guesses change if we
modify our pre-data assumptions? We show, in the same section, that the
typical features indicated by our method are robust against such changes.

@@

% (for a full network of 1\,000 neurons and
% a recording of 500\,000 time bins, for example, there are
% $\binom{500\,000+1\,000}{1\,000} \approx 10^{3\,131}$ possible frequency
% distributions).

@@

% \mynote{Add this?: from sampling theory we know that important
%   features of the full network may not be visible in a sample because
%   smoothed out. But using sampling theory in the inverse direction we can
%   infer such full-network features from the sample.}

@@

The context we consider is as follows. During an experimental session we
have recorded the spiking activities of $n$ neurons for a certain amount of
time. These neurons are our \enquote{sample} or \enquote{subnetwork}. Their
spikes are binned into $T$ time bins and binarized to $\set{0,1}$ values in
each bin. Call $\ya_{t}$ the number of neurons that fire during time bin
$t$: this is the \emph{total activity} of the sample, or just
\enquote{activity} for short. Obviously $\ya_{t} \in\set{0,1,\dotsc,n}$; if
$\ya_{t}=0$, no neuron spikes during bin $t$; if $\ya_{t}=n$, all spike at
some point during bin $t$. For brevity, let's say \enquote{at $t$} for
\enquote{during time bin $t$}. If we divide the total activity by the
population size we have the normalized total activity or
population-averaged activity $\ya/n$, ranging from $0$ to $1$ in $1/n$
steps. From the activities $\set{\ya_{t}}$ we can count how often the
activity levels $\ya=0$, $\ya=1$, and so on appeared during the recording,
obtaining the distribution of measured relative frequencies
$(\yff_{\ya}) \defs \yf$. We can also consider the sample activity at time
bins \emph{outside} of the recorded period. Such activity is unknown to us,
of course.

For many animal species, the neurons that are recorded within a brain area
are not specifically chosen from among the rest, owing to several limiting
factors; for example, limitations in how precisely electrodes are inserted
or neurons are targeted by viruses. The set of recorded neurons may even
change slightly across experimental sessions that are very far apart in
time. We assume that there's an area, comprising a network of $N$ neurons,
for which we believe that any other sample of size $n$ could have equally
plausibly been recorded instead of the sample of $n$ neurons that was
actually recorded. We call this larger network the \enquote{full network}.
The total activity of these $N$ neurons at $t$ is $\yA_{t}$. The relative
frequencies of the various activity levels during the recording were
$(\yFF_{\yA}) \defs \yF$. We don't know the values $\yA_{t}$ at each $t$,
or the frequency distribution $\yF$. We only know for certain that
$\yA_{t} \in \set{0, 1, \dotsc, N}$ and that $\yA_{t} \ge \ya_{t}$ for
obvious reasons. For the time being we assume that we know $N$; in
\sect~\ref{sec:N} we discuss the consequences of our lack of precise
knowledge about this number.

Our questions concern general features of the total activity $\yA$ of the
full network during and after the recording, and across sessions under the
same study conditions. For example: what was its frequency distribution
during the recording? How much does this frequency distribution change
across sessions? How much total activity should we expect at any time bin
during a recording? The approach presented here gives a probability
distribution that approximately answers all these questions.

@@

The idea behind our approach is easily summarized:
\begin{enumerate}[label=(\alph*)]
\item\label{item:sample_step} Using sampling theory we determine the
  relation between some expected values -- specifically, moments -- for the
  total activity $\ya$ of the sample and corresponding expected values $M$
  for the total activity $\yA$ of the full network.
\item\label{item:maxent_step} Using the maximum-entropy method we build a
  distribution $\px(\yA \| M, N)$ for the total activity of the full
  network of $N$ neurons, using as constraints the expected values $M$
  found in the previous step.
\end{enumerate}
We now discuss the ideas behind these steps more in detail, leaving their
precise mathematical implementation and a more detailed list of references
to appendix~\ref{sec:derivation_maxent_fullnet}.

@@

Step~\ref{item:sample_step}, a result of sampling theory, is just an
application of the probability calculus, which gives an exact linear
relation between the first $m$ moments for the full network and the first
$m$ for the sample \citep[\eqns~(16)]{portamanaetal2015}. The ones
determine the others and vice versa at every time bin. This relation holds
for any belief distribution $\pxxx(\yA_{t})$ for the full-network activity
and its marginal $\pxx(\ya_{t})$ for the sample activity at that bin.

This sampling relation is even more straightforward if instead of power
moments we use \emph{normalized factorial moments} \citep{potts1953}. The
$m$th normalized factorial moment of a distribution $\pxx(\ya)$ is
\begin{equation}
  \label{eq:def_factorial_moment_a}
  \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
  \defd
  \sum_{\ya=0}^{n} %m!\,
  \displaystyle\binom{\ya}{m} \bigg/ \binom{n}{m}\; \pxx(\ya),
  \qquad 1\le m \le n,
\end{equation}
% where $\px(\ya \| \yH)$ is the probability for activity $\ya$ given
% assumptions $\yH$; \enquote{$\ya$} may refer to any recorded bin or to a
% new bin.
that is, the expected number of distinct $m$-tuples of simultaneously
spiking neurons (within a time-bin's width), normalized by the maximum
possible number of distinct $m$-tuples. % For
% example, with $m=2$, if $a=4$ neurons spike in a network of $n=5$, we
% have $\tbinom{4}{2}=6$ distinct pairs of simultaneously spiking neurons,
% and the total number of distinct pairs is $\tbinom{5}{2}=10$. The
% normalized number of spiking pairs is therefore $6/10$.
Note that the first $m$ factorial moments together provide the same
information as the first $m$ power moments together, and vice versa: they
are linearly related because $\tbinom{\ya}{m}$ is a polynomial in $\ya$ of
degree $m$. % Specifying one set is therefore
% equivalent to specifying the other set.
But the normalized factorial moments have a convenient property for our
analysis: \emph{the first $n$ normalized factorial moments for the sample
  and for the full network are numerically identical}:
\begin{equation}
  \label{eq:equal_m_factorialmoments}
  \begin{split}
  \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
  &=
  \E\Biggl[\binom{\yA}{m} \Biggr]\bigg/\binom{N}{m}
  \qquad\text{or}\\
    \sum_{\ya=0}^{n} %m!\,
  \binom{\ya}{m}\bigg/  \binom{n}{m}\; \pxx(\ya)
&=    \sum_{\yA=0}^{N} %m!\,
  \binom{\yA}{m}\bigg/  \binom{N}{m}\; \pxxx(\yA),\qquad
  % \E\condt[\big]{\tbinom{\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
  % \E\condt[\big]{\tbinom{\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
  1\le m \le n.
\end{split}
\end{equation}
% where $\pxxx(\yA)$ is our distribution of belief about the full-network
% activity, and $\pxx$

@@

In our case we don't have measured time averages of the quantity
$\binom{\yA}{m}$ to equate to the factorial moment
$\E\bigl[\binom{\yA}{m}\bigr]$ of the full-network distribution
$\pxxx(\yA)$. But \eqn~\eqref{eq:equal_m_factorialmoments} of
step~\ref{item:sample_step} comes to our rescue, because it says that the
expectations for the full network are determined by those for the sample
$\E\bigl[\binom{\ya}{m}\bigr]$, and we do have the time averages of the
corresponding sample quantity $\binom{\ya}{m}$. So we can combine the two
steps:
\begin{equation*}
  \label{eq:chain_prescription_moments}
  \mathrlap{\overbracket{\phantom{\text{measured moments} \rightarrow
        \text{sample moments}}}^{\textit{maximum-entropy prescription}}}
  \text{measured moments} \rightarrow
        \underbracket{\text{sample moments}
  \rightarrow
    \text{full-network moments}}_{\textit{sampling theory}}
\end{equation*}
and apply the maximum-entropy method to obtain a distribution
$\px(\yA \| M, N)$ for the full network of $N$ neurons by constraining some
of its factorial moments, for example $M=\set{1,\dotsc,m'}$ with $m'\le n$,
to be equal to the sample's recorded averages. In formulae, the constraints
on $\pxxx(\yA)$ are
\begin{equation}
  \label{eq:final_constraint}
\underbracket{\frac{1}{T}\sum_{t}\binom{\ya_{t}}{m}\bigg/\binom{n}{m}
  \equiv
   \sum_{\ya}
  \binom{\ya}{m}\bigg/  \binom{n}{m} \; \yff_{\ya}}_{\text{measured moments}}
  \mathrel{\;=\;}
   \underbracket{\sum_{\yA}
     \binom{\yA}{m}\bigg/  \binom{N}{m} \; \pxxx(\yA)}_{\text{distribution moments}},
   \quad m\in M.
\end{equation}

@@

We apply the method just described to two data sets publicly available in
the literature:
\begin{itemize}[wide]
\item The \textbf{first data set}, from \textcite[rat
  14147]{stensolaetal2012}, consists of $n=65$ neurons (27 of which
  classified as grid cells) from rat Medial Entorhinal Cortex, recorded for
  about 20 minutes. Their spikes are binned into $T=417\,641$ bins of
  $3\;\textrm{ms}$ width.
\item The \textbf{second data set}, from
  \textcite{rostamietal2016_r2017}, %\texttt{data\_RS.npy}
  consists of $n=159$ neurons from macaque Motor Cortex, recorded for about
  15 minutes. Their spikes are binned into $T=300\,394$ bins of
  $3\;\textrm{ms}$ width.
\end{itemize}

% In the followingFrom a Bayesian point of view, the maximum-entropy distribution $\px(\yA)$
% can be considered as \emph{the most probable frequency distribution} for
% the full-network activity during the recording. But it also approximates
% three other distributions, as discussed in the previous section.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-25T17:35:14+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Abstract
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abstractrunin
\abslabeldelim{}
\renewcommand*{\abstractname}{}
\setlength{\absleftindent}{0pt}
\setlength{\absrightindent}{0pt}
\setlength{\abstitleskip}{-\absparindent}
\begin{abstract}\labelsep 0pt%
  \noindent \mynote{Abstract must be rewritten once paper is ready}This
  work shows how to build a maximum-entropy probabilistic model for the
  total activity of a population of neurons, given only some activity data or
  statistics -- for example, empirical moments -- of a \emph{subpopulation}
  thereof. This kind of model is useful because neuronal recordings are
  always limited to a very small sample of a population of neurons.
%
  The model is applied to two sets of neuronal data available in the
  literature. In some cases it makes interesting forecasts about the larger
  population -- for example, two low-regime modes in the frequency
  distribution for the total activity -- that are not visible in the sample
  data or in maximum-entropy models applied only to the sample.
%
  For the two datasets, the maximum-entropy probability model applied only
  to the subpopulation is compared with the marginal probability
  distribution obtained from the maximum-entropy model applied to the full
  population. On a linear probability scale no large differences are
  visible, but on a logarithmic scale the two distributions show very
  different behaviours, especially in the tails.
  \\\noindent\emph{\footnotesize Note: Dear Reader \amp\ Peer, this
    manuscript is being peer-reviewed by you. Thank you.}
% \par%\\[\jot]
% \noindent
% {\footnotesize PACS: ***}\qquad%
% {\footnotesize MSC: ***}%
%\qquad{\footnotesize Keywords: ***}
\end{abstract}
\selectlanguage{british}\frenchspacing

@@

\section{Introduction:\\a simple model for questions about large neuronal
  populations}
\label{sec:intro}

@@

Despite remarkable advances in recording technologies, the best
experimental measurements of instantaneous neuronal activity can still only
record a very small sample of neurons -- hundreds at most -- compared to
the numbers that constitute a functionally distinguished brain region. Many
probabilistic models focus on such samples only, somehow neglecting, in
their assumptions, that the recorded neurons are a sample from a larger
population. This kind of isolation assumptions sometimes escape attention,
being subtly hidden in the mathematics. Some probabilistic models try to
take also unrecorded neurons into account individually, and therefore
become very complex\mynote{Yasser \amp\ Vahid: refs for here?}. It would be
useful to have models that operate in between: addressing the larger brain
area whence the sample comes, but without adding too much complexity and
detail about it. Such intermediate models would be useful, for example, for
preliminary investigations, to help us to decide which hypotheses to
discard and which to consider for more complex and costly studies.

\mynote{[Yasser:] In the present work we aim to address the issue of building probability distributions over a large population using recording from subpopulations by considering the distribution of total population activity as an example. [}
In the present work we propose such an intermediate model. It answers this
question:\mynote{]} How much can the \emph{total} activity of a large neuronal
population have been, given the observation of the activity of a very small
sample thereof? This model addresses a larger brain area, avoiding the
assumption of isolation of the sample; and by focusing on the total
activity, rather than the activity of individual neurons, it remains simple
and numerically tractable.

@@

We illustrate possible uses of our proposed model in
\sect~\ref{sec:application}, by applying it to two concrete data sets:
\begin{enumerate*}[label=(\alph*)]
\item the activity of 65 neurons recorded for 20 min from a rat's Medial
  Entorhinal Cortex \citep{stensolaetal2012}, \item the activity of 159
  neurons recorded for 15 min from a macaque's Motor Cortex
  \citep{rostamietal2016_r2017}.
\end{enumerate*}
For each data set the model gives the most plausible frequency distribution
with which all levels of total activity of a larger population of 1\,000--10\,000
neurons appeared during the recording. These guessed frequency
distributions are distinctly different from the recorded ones for the
sampled neurons. For the first data set, for example, the frequency
distribution for the full population has two very distinct modes, both at low
activities, %($0.7\%$ and $2.6\%$ of simultaneously active neurons
% within $3\;\mathrm{ms}$,
see \fig~\ref{fig:plots_5mom}, whereas the frequency distribution for the
sample is monotonically decreasing with its maximum at zero activity. The
frequency distribution for the second data set doesn't have two modes but
still presents one prominent shoulder in its low-activity mode. Note that
these guessed features of the full population cannot be inferred by the
application of maximum-entropy \emph{to the sample alone}. Although we don't
formulate any hypotheses on why the larger population could have two distinct
low-activity regimes -- the purpose of the present work is only
methodological -- these results shows that the proposed method can lead to
the formulation or preliminary assessment of interesting hypotheses.

We want to stress the usefulness of making a quantified guess about the
activity of a larger brain area. Such a guess seems indeed to be the
primary idea behind recording a sample from that area. %  Second, it allows us
% to make comparisons across experimental sessions; such comparisons would me
% difficult or meaningless if made with the recorded samples, which generally
% comprise non-overlapping sets of neurons and differ in size.
There's also an important methodological reason. Maximum-entropy methods
have been used to assess the informational sufficiency of pairwise
correlations and correlations of higher order \citep[see for
example][]{martignonetal1995,bohteetal2000,schneidmanetal2006,shlensetal2006,barreiroetal2010,ganmoretal2011,granotatedgietal2013}.
But when the difference in size between sample and full population is too
large, correlation sufficiency for the sample implies the \emph{lack} of
correlation sufficiency for the larger population, and vice versa. Therefore,
maximum-entropy applications at the sample level can be deceptive for
questions of statistical sufficiency, and a full-population application is
more reliable. We discuss this point in detail in \sect~\ref{sec:marginalization}.

@@

How large is the full population addressed by the method proposed here? Its
size must have some limit and can't obviously include the full brain. The
size is determined by the validity of the formulae from sampling theory and
discussed in \sect~\ref{sec:N}.

We obviously don't know whether the frequency distribution obtained with
our approach is the \emph{actual} one which the activity levels of the full
population had during the recording; it's only the most plausible. In high
dimensions, however, the features of the most plausible distribution may
\emph{not} be typical of the majority of most plausible distributions; and
the set of all possible frequency distributions, if the full population for
example comprises 1\,000 neurons, is a 1\,000-dimensional space. In
\sect~\ref{sec:assumptions} we therefore try to assess which features of
the frequency distribution delivered by our method may be typical and
therefore expected of the actual one. We find that general features such as
the bimodality of the first data set are indeed typical. Maximum-entropy
models can be considered as approximations of Bayesian models based on
various assumptions of inferential sufficiency
\citep{jaynes1986d_r1996,portamana2017}. How do our guesses change if we
modify our pre-data assumptions? We show, in the same section, that the
typical features indicated by our method are robust against such changes.

@@

% (for a full population of 1\,000 neurons and
% a recording of 500\,000 time bins, for example, there are
% $\binom{500\,000+1\,000}{1\,000} \approx 10^{3\,131}$ possible frequency
% distributions).

@@

% \mynote{Add this?: from sampling theory we know that important
%   features of the full population may not be visible in a sample because
%   smoothed out. But using sampling theory in the inverse direction we can
%   infer such full-population features from the sample.}

@@

The context we consider is as follows. During an experimental session we
have recorded the spiking activities of $n$ neurons for a certain amount of
time. These neurons are our \enquote{sample} or \enquote{subpopulation}. Their
spikes are binned into $T$ time bins and binarized to $\set{0,1}$ values in
each bin. Call $\ya_{t}$ the number of neurons that fire during time bin
$t$: this is the \emph{total activity} of the sample, or just
\enquote{activity} for short. Obviously $\ya_{t} \in\set{0,1,\dotsc,n}$; if
$\ya_{t}=0$, no neuron spikes during bin $t$; if $\ya_{t}=n$, all spike at
some point during bin $t$. For brevity, let's say \enquote{at $t$} for
\enquote{during time bin $t$}. If we divide the total activity by the
population size we have the normalized total activity or
population-averaged activity $\ya/n$, ranging from $0$ to $1$ in $1/n$
steps. From the activities $\set{\ya_{t}}$ we can count how often the
activity levels $\ya=0$, $\ya=1$, and so on appeared during the recording,
obtaining the distribution of measured relative frequencies
$(\yff_{\ya}) \defs \yf$. We can also consider the sample activity at time
bins \emph{outside} of the recorded period. Such activity is unknown to us,
of course.

For many animal species, the neurons that are recorded within a brain area
are not specifically chosen from among the rest, owing to several limiting
factors; for example, limitations in how precisely electrodes are inserted
or neurons are targeted by viruses. The set of recorded neurons may even
change slightly across experimental sessions that are very far apart in
time. We assume that there's an area, comprising a population of $N$ neurons,
for which we believe that any other sample of size $n$ could have equally
plausibly been recorded instead of the sample of $n$ neurons that was
actually recorded. We call this larger population the \enquote{full population}.
The total activity of these $N$ neurons at $t$ is $\yA_{t}$. The relative
frequencies of the various activity levels during the recording were
$(\yFF_{\yA}) \defs \yF$. We don't know the values $\yA_{t}$ at each $t$,
or the frequency distribution $\yF$. We only know for certain that
$\yA_{t} \in \set{0, 1, \dotsc, N}$ and that $\yA_{t} \ge \ya_{t}$ for
obvious reasons. For the time being we assume that we know $N$; in
\sect~\ref{sec:N} we discuss the consequences of our lack of precise
knowledge about this number.

Our questions concern general features of the total activity $\yA$ of the
full population during and after the recording, and across sessions under the
same study conditions. For example: what was its frequency distribution
during the recording? How much does this frequency distribution change
across sessions? How much total activity should we expect at any time bin
during a recording? The approach presented here gives a probability
distribution that approximately answers all these questions.

@@

The idea behind our approach is easily summarized:
\begin{enumerate}[label=(\alph*)]
\item\label{item:sample_step} Using sampling theory we determine the
  relation between some expected values -- specifically, moments -- for the
  total activity $\ya$ of the sample and corresponding expected values $M$
  for the total activity $\yA$ of the full population.
\item\label{item:maxent_step} Using the maximum-entropy method we build a
  distribution $\px(\yA \| M, N)$ for the total activity of the full
  population of $N$ neurons, using as constraints the expected values $M$
  found in the previous step.
\end{enumerate}
We now discuss the ideas behind these steps more in detail, leaving their
precise mathematical implementation and a more detailed list of references
to appendix~\ref{sec:derivation_maxent_fullnet}.

@@

Step~\ref{item:sample_step}, a result of sampling theory, is just an
application of the probability calculus, which gives an exact linear
relation between the first $m$ moments for the full population and the first
$m$ for the sample \citep[\eqns~(16)]{portamanaetal2015}. The ones
determine the others and vice versa at every time bin. This relation holds
for any belief distribution $\pxxx(\yA_{t})$ for the full-population activity
and its marginal $\pxx(\ya_{t})$ for the sample activity at that bin.

This sampling relation is even more straightforward if instead of power
moments we use \emph{normalized factorial moments} \citep{potts1953}. The
$m$th normalized factorial moment of a distribution $\pxx(\ya)$ is
\begin{equation}
  \label{eq:def_factorial_moment_a}
  \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
  \defd
  \sum_{\ya=0}^{n} %m!\,
  \displaystyle\binom{\ya}{m} \bigg/ \binom{n}{m}\; \pxx(\ya),
  \qquad 1\le m \le n,
\end{equation}
% where $\px(\ya \| \yH)$ is the probability for activity $\ya$ given
% assumptions $\yH$; \enquote{$\ya$} may refer to any recorded bin or to a
% new bin.
that is, the expected number of distinct $m$-tuples of simultaneously
spiking neurons (within a time-bin's width), normalized by the maximum
possible number of distinct $m$-tuples. % For
% example, with $m=2$, if $a=4$ neurons spike in a population of $n=5$, we
% have $\tbinom{4}{2}=6$ distinct pairs of simultaneously spiking neurons,
% and the total number of distinct pairs is $\tbinom{5}{2}=10$. The
% normalized number of spiking pairs is therefore $6/10$.
Note that the first $m$ factorial moments together provide the same
information as the first $m$ power moments together, and vice versa: they
are linearly related because $\tbinom{\ya}{m}$ is a polynomial in $\ya$ of
degree $m$. So we'll just say \enquote{first $m$ moments} from now
on. % Specifying one set is therefore
% equivalent to specifying the other set.
But the normalized factorial moments have a convenient property for our
analysis: \emph{the first $n$ normalized factorial moments for the sample
  and for the full population are numerically identical}:
\begin{equation}
  \label{eq:equal_m_factorialmoments}
  \begin{split}
  \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
  &=
  \E\Biggl[\binom{\yA}{m} \Biggr]\bigg/\binom{N}{m}
  \qquad\text{or}\\
    \sum_{\ya=0}^{n} %m!\,
  \binom{\ya}{m}\bigg/  \binom{n}{m}\; \pxx(\ya)
&=    \sum_{\yA=0}^{N} %m!\,
  \binom{\yA}{m}\bigg/  \binom{N}{m}\; \pxxx(\yA),\qquad
  % \E\condt[\big]{\tbinom{\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
  % \E\condt[\big]{\tbinom{\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
  1\le m \le n.
\end{split}
\end{equation}
% where $\pxxx(\yA)$ is our distribution of belief about the full-population
% activity, and $\pxx$

@@

In our case we don't have measured time averages of the quantity
$\binom{\yA}{m}$ to equate to the factorial moment
$\E\bigl[\binom{\yA}{m}\bigr]$ of the full-population distribution
$\pxxx(\yA)$. But \eqn~\eqref{eq:equal_m_factorialmoments} of
step~\ref{item:sample_step} comes to our rescue, because it says that the
expectations for the full population are determined by those for the sample
$\E\bigl[\binom{\ya}{m}\bigr]$, and we do have the time averages of the
corresponding sample quantity $\binom{\ya}{m}$. So we can combine the two
steps:
\begin{equation*}
  \label{eq:chain_prescription_moments}
  \mathrlap{\overbracket{\phantom{\text{measured moments} \rightarrow
        \text{sample moments}}}^{\textit{maximum-entropy prescription}}}
  \text{measured moments} \rightarrow
        \underbracket{\text{sample moments}
  \rightarrow
    \text{full-population moments}}_{\textit{sampling theory}}
\end{equation*}
and apply the maximum-entropy method to obtain a distribution
$\px(\yA \| M, N)$ for the full population of $N$ neurons by constraining some
of its factorial moments, for example $M=\set{1,\dotsc,m'}$ with $m'\le n$,
to be equal to the sample's recorded averages. In formulae, the constraints
on $\pxxx(\yA)$ are
\begin{equation}
  \label{eq:final_constraint}
\underbracket{\frac{1}{T}\sum_{t}\binom{\ya_{t}}{m}\bigg/\binom{n}{m}
  \equiv
   \sum_{\ya}
  \binom{\ya}{m}\bigg/  \binom{n}{m} \; \yff_{\ya}}_{\text{measured moments}}
  \mathrel{\;=\;}
   \underbracket{\sum_{\yA}
     \binom{\yA}{m}\bigg/  \binom{N}{m} \; \pxxx(\yA)}_{\text{distribution moments}},
   \quad m\in M.
\end{equation}

@@

We apply the method just described to two data sets publicly available in
the literature:
\begin{itemize}[wide]
\item The \textbf{first data set}, from \textcite[rat
  14147]{stensolaetal2012}, consists of $n=65$ neurons (27 of which
  classified as grid cells) from rat Medial Entorhinal Cortex, recorded for
  about 20 minutes. Their spikes are binned into $T=417\,641$ bins of
  $3\;\textrm{ms}$ width.
\item The \textbf{second data set}, from \textcite[data courtesy by A.
  Riehle and T. Brochier]{rostamietal2016_r2017}, %\texttt{data\_RS.npy}
  consists of $n=159$ neurons from macaque Motor Cortex, recorded for about
  15 minutes. Their spikes are binned into $T=300\,394$ bins of
  $3\;\textrm{ms}$ width.
\end{itemize}
